Synthesizing Symmetric Lenses

### 2018-07-12 papers/SymmetricLenses/main.tex
pass through evaluation

To resolve this tension, our algorithm uses heuristics to judge whether to
return the current best satisfying lens to the user or to pass the
next set of unrollings to \GreedySynth.
The heuristics favor stopping if the current best satisfying
lens has relatively high likelihood, indicating the lens is very promising.
The heuristics also favor stopping if \Expand
has delievered to \GreedySynth all pairs of stochastic regular
expressions at a given rank and there is a large number of pairs at the next
rank because searching through all such pairs will take a long time.
With this approach, the algorithm can quickly respond with a satisfying
lens with relatively high likelihood.  If users are unhappy with the result, they can refine the
search by supplying additional examples, which serve both to rule out
the previously proposed lens(es) and to reduce the size of the search
space by cutting down on the number of satisfying lenses.

@@

\subsection{Optimizations}
\label{subsec:optimizations}

Our implementation includes a number of optimizations not described above:
annotations that guide DNF conversion; an expansion inference
algorithm; hash-consing and memoization; and compositional synthesis. The
implementation also provides for an additional specification parameter that
customizes the termination condition for the synthesis procedure.

@@

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented symmetric DNF lenses as an extension to
Boomerang. Specifically, we added symmetric lens combinators to
Boomerang, and implemented all of Boomerang's asymmetric lens
combinators  using a combination of the symmetric lens combinators presented in this
paper and symmetric versions of the asymmetric extensions
already present in Boomerang, like matching lenses and quotient lenses.
We also integrated our synthesis engine into Boomerang. This
integration allows users to write synthesis tasks 
alongside lens combinators, incorporate synthesis results into 
manually-written lenses, and reference previously defined lenses in
the synthesis of new lenses.

In this evaluation, we aim to answer four primary questions:
\begin{enumerate}
\item Does the synthesis algorithm find the correct lens?
\item Is the synthesis procedure efficient enough to be used in everyday
  development?
\item How much slower is our tool on bijective lens synthesis benchmarks than
  the existing Optician system customized for bijective lenses?

\item Which heuristics are useful
for reducing the ``specification 
  overhead'' for users? In particular, which heuristics are useful in
  synthesizing the desired lens instead of merely a lens that satisfies the
  specification?
\end{enumerate}

\subsection{Benchmark Suite}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted 8 data cleaning benchmarks from Flash Fill. Note
that our tool produces bidirectional transformations rather
than one-way transformers like Flash Fill. 
\item We adapted 29 benchmarks from Augeas. These benchmarks turn the
  ad hoc data present in Linux configuration files into a structured dictionary
  representation.
\item We created 11 additional benchmarks 
  derived from real-world examples and/or the bidirectional programming
  literature. These tasks range from synchronizing REST and JSON web resource
  descriptions to synchronizing BibTeX and EndNote citation descriptions.
\end{enumerate}
Each benchmark comprises a source and target regular expression, a
collection of examples, and an optional termination condition
parameter. 
%%A realistic benchmarks suite is critical in answering the first research
%%question. To determine whether our algorithm is fast enough to be used in
%%everyday development, we need to evaluate whether we can solve real-world
%%problems efficiently.

All three sources for our benchmarks provide realistic problems
suitable for answering the research questions:
Flash Fill  data cleaning tasks are either derived from online help
forums or taken from the Excel product team;  
Augeas is a utility that transforms Linux configuration files into a manageable dictionary
form; and
many of the remaining benchmarks arise from synchronizing real-world
data formats such as BibTeX and EndNote.

\subsection{Lens correctness}
To determine whether the synthesized lenses were the desired ones,
we ran the symmetric synthesis algorithm on the 48 benchmarks and then
manually inspected the generated lenses.  In all 48 cases, the
synthesized benchmark was correct.

\subsection{Usefulness in Everyday Development}

We measured the run time of our algorithm in two modes:

\begin{tabulary}{\linewidth}{rL}
  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
  \SSNCOpt{}: & The symmetric synthesis algorithm, but compositional synthesis not enabled.\\
\end{tabulary}\\

Recall that compositional synthesis exploits user annotations to break a
benchmark into a series of smaller synthesis steps. In our experience,
compositional synthesis (SS mode) allows our system to scale to arbitrarily
large and complex formats. However, SSNC mode, which uses no such annotations
and synthesizes a complete lens all at once provides a useful experimental
stress test for the system.

\begin{figure}
  \includegraphics{generated-graphs/times}
  \caption{Number of benchmarks that can be solved by a given algorithm in a
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \SSNCOpt{} is the symmetric synthesis algorithm without using a library of
    existing lenses. The symmetric synthesis algorithm is able to complete all
    benchmarks in under 30 seconds elapsed total time. In addition, it is able to
    complete a majority of the benchmarks in that timeframe without
  using compositional 
    synthesis.}
  \label{fig:times}
\end{figure}

For each benchmark in our benchmark suite, we ran it in both modes with a
timeout of 60 seconds, and averaged the result over 5 runs.
Figure~\ref{fig:times} summarizes the results of these tests. We find that our
algorithm is able to synthesize all of the benchmarks in our benchmark suite in
under 30 seconds. Without compositional synthesis, the synthesis algorithm is
able to solve 26 problem instances. Recall from
Section~\ref{subsec:optimizations} that users may tune the termination metric.
Success on this set of benchmarks requires we tune that termination condition
(e.g. for some benchmarks, allowing the system flexibility to spend more time
considering more of the search space). In
Section~\ref{subsec:effect-of-heuristics}, we will consider the effect of fixing
the termination metric. In practice, we found customizing the termination metric
to be quite easy: if manual inspection of the lens showed there were too few
bijections, we would increase the termination parameter. If synthesis took too
long, we would decrease it.

@@

To compare to the existing bijective synthesis algorithm, we run our symmetric
synthesis algorithm on the original Optician benchmarks, comprised of
39 bijective synthesis tasks. We had to slightly alter four of these
benchmarks, either by providing additional examples or by tuning termination parameters.
Without these alterations, symmetric synthesis yielded a lens that fit
the specification but that was not the desired lens.

@@

Figure~\ref{fig:metric}
summarizes the result of these experiments. The data reveal that 
the information-theoretic metric is critical for finding the correct
lens: Only 10 of the benchmarks succeeded when running in \CCOpt{} mode.
The termination condition is also quite important. When running
in \FLOpt{} mode, the algorithm only discovers 5 lenses, which shows
that the first cluster that contains a satisfying lens is rarely the
correct cluster.  However, our termination condition is not perfect.
Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
Without termination parameters, our algorithm finds the correct lens for 36 of
our 48 benchmarks.  Nine of them need help in the form of a termination parameter
for finding the correct lens. That being said, our default parameter finds a
good middle-ground. Both \TFTOpt{} and \NTFTOpt{} synthesized the correct lens
less often than \ZTOpt.  \TFTOpt{} demands too good a lens, so the search
procedure keeps spinning, while \NTFTOpt{} terminates early, not
always trying hard enough to find the correct lens.
% end evaluation

>>>>

To resolve this tension, our algorithm uses heuristics to judge whether to
return the current best satisfying lens to the user or to pass the next set of
unrollings to \GreedySynth.  The heuristics favor stopping if the current best
satisfying lens has relatively high likelihood, indicating the lens is very
promising.  The heuristics also favor stopping if \Expand has delivered to
\GreedySynth all pairs of stochastic regular expressions at a given rank and
there is a large number of pairs at the next rank because searching through all
such pairs will take a long time.  With this approach, the algorithm can quickly
respond with a satisfying lens with relatively high likelihood.  If users are
unhappy with the result, they can either refine the search by supplying
additional examples, which serve both to rule out the previously proposed
lens(es) and to reduce the size of the search space by cutting down on the
number of satisfying lenses, or supply a tuning parameter (discussed in 
\S\ref{subsec:optimizations}) that adjusts how much effort is put into the search.

@@

\subsection{Optimizations and Termination Tuning}
\label{subsec:optimizations}

Our implementation includes a number of optimizations not described above:
annotations that guide DNF conversion; an expansion inference
algorithm; hash-consing and memoization; and compositional synthesis. The
implementation also provides for an additional specification parameter that
customizes the termination condition for the synthesis procedure.  The
optimizations make the system performant enough for interactive use, and the
tuning parameter gives the user a ``knob'' by which the algorithm can be
directed to try harder to find a good lens.

@@

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented symmetric DNF lenses as an extension to
Boomerang. Specifically, we added symmetric lens combinators to
Boomerang, and implemented all of Boomerang's asymmetric lens
combinators  using a combination of the symmetric lens combinators presented in this
paper and symmetric versions of the asymmetric extensions (like matching lenses and quotient lenses)
already present in Boomerang.
We also integrated our synthesis engine into Boomerang. This
integration allows users to write synthesis tasks 
alongside lens combinators, incorporate synthesis results into 
manually-written lenses, and reference previously defined lenses in
the synthesis of new lenses.

In this evaluation, we aim to answer four primary questions:
\begin{enumerate}
\item Can the algorithm (with suitable examples and termination
  tuning) find the correct lens?
\item Is the synthesis procedure efficient enough to be used in everyday
  development?
\item How much slower is our tool on bijective lens synthesis benchmarks than
  the existing Optician system customized for bijective lenses?
\item How much termination tuning is necessary and how does the heuristic affect the
  results?
\end{enumerate}

\subsection{Benchmark Suite}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted 8 data cleaning benchmarks from Flash Fill. Note
that our tool produces bidirectional transformations rather
than one-way transformers like Flash Fill. 
\item We adapted 29 benchmarks from Augeas. These benchmarks turn the
  ad hoc data present in Linux configuration files into a structured dictionary
  representation.
\item We created 11 additional benchmarks 
  derived from real-world examples and/or the bidirectional programming
  literature. These tasks range from synchronizing REST and JSON web resource
  descriptions to synchronizing BibTeX and EndNote citation descriptions.
\end{enumerate}

All three sources for our benchmarks provide realistic problems suitable for
answering the research questions: Flash Fill data cleaning tasks are either
derived from online help forums or taken from the Excel product team; Augeas is
a utility that transforms Linux configuration files into a manageable dictionary
form; and many of the remaining benchmarks arise from synchronizing real-world
data formats such as BibTeX and EndNote.  Each task consists of a source and
target regular expression.\saz{Do we want to say anything about the sizes of these?}

\subsection{Synthesizing Correct Lenses}

Because our tool is intended to be interactive---the user typically starts with
zero or one examples, runs the tools, determines whether the output is the
desired one and, if not, either adds an example or adjusts the termination
metric and then tries again---it is hard to cleanly evaluate the net user
experience.  The primary concern is whether it is usable to generate
synthesize correct lenses.

To determine whether the system can synthesize the desired lenses, we ran the
symmetric synthesis algorithm interactively on all 48 tasks,
working with the system to create sufficient examples and adjust the termination
tuning.  In all cases, the desired lens was obtained, which we verified by
manually inspecting the generated lens programs.
%
The majority of the tasks required only a single example and none required
more than three examples in order to synthesize the desired lens.\footnote{In
  one benchmark, we supplied a fourth example that was later discovered to 
  be unnecessary.}

Adjusting the tuning parameter was needed in only 9 of the 48 tasks. In practice, we
found that customizing the termination metric to be quite easy: if manual
inspection of the lens showed there were too few bijections, we would increase
the termination parameter. If synthesis took too long, we would decrease it.
Section~\ref{subsec:effect-of-heuristics}, studies the effect of fixing the
termination metric.

%%A realistic benchmarks suite is critical in answering the first research
%%question. To determine whether our algorithm is fast enough to be used in
%%everyday development, we need to evaluate whether we can solve real-world
%%problems efficiently.

\subsection{Effectiveness of Compositional Synthesis}

Having determined appropriate examples and timing parameters for the 48
benchmarks, we evaluate the performance of the system by 
measuring the running time of our algorithm in two modes:

\begin{tabulary}{\linewidth}{rL}
  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
  \SSNCOpt{}: & The symmetric synthesis algorithm, with no compositional synthesis enabled.\\
\end{tabulary}\\

Recall that compositional synthesis exploits user annotations to break a
benchmark into a series of smaller synthesis steps. In our experience,
compositional synthesis (SS mode) allows our system to scale to arbitrarily
large and complex formats; measuring it provides an estimate of the
responsiveness of the system when used as intended.\saz{Anders check this!} However, SSNC mode, which
uses no such annotations and synthesizes a complete lens all at once provides a
useful experimental stress test for the system.

\begin{figure}
  \includegraphics{generated-graphs/times}
  \caption{Number of benchmarks that can be solved by a given algorithm in a
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \SSNCOpt{} is the symmetric synthesis algorithm without using a library of
    existing lenses. The symmetric synthesis algorithm is able to complete all
    benchmarks in under 30 seconds elapsed total time. Without compositional
    synthesis it is able to complete 26. Each benchmark specification includes
    source and target regular expressions, between one and three sufficient
    examples, and, if necessary, the timing parameter determined to be
    sufficient for correctness during a prior interactive session.}
  \label{fig:times}
\end{figure}

For each benchmark in the suite and each mode, we ran the system with a timeout
of 60 seconds, averaging the result over 5 runs.  Figure~\ref{fig:times}
summarizes the results of these tests. We find that our algorithm is able to
synthesize all of the benchmarks in under 30 seconds. Without compositional
synthesis, the synthesis algorithm is able to solve only 26 problem
instances.

% Section~\ref{subsec:optimizations} that users may tune the termination metric.
% Success on this set of benchmarks requires we tune that termination condition
% (e.g. for some benchmarks, allowing the system flexibility to spend more time
% considering more of the search space). In

@@

To compare to the existing bijective synthesis algorithm, we run our symmetric
synthesis algorithm on the original Optician benchmarks, comprised of
39 bijective synthesis tasks.\footnote{We had to slightly alter four of these
benchmarks, either by providing additional examples or by tuning termination parameters.
Without these alterations, symmetric synthesis yielded a lens that fit
the specification but that was undesired.}

@@

Figure~\ref{fig:metric}
summarizes the result of these experiments. The data reveal that 
the information-theoretic metric is critical for finding the correct
lens: Only 10 of the benchmarks succeeded when running in \CCOpt{} mode.
The termination condition is also quite important. When running
in \FLOpt{} mode, the algorithm only discovers 5 lenses, which shows
that the first cluster that contains a satisfying lens is rarely the
correct cluster.  However, our termination condition is not perfect.
Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
Without termination parameters, our algorithm finds the correct lens for 39 of
our 48 benchmarks; nine required tuning 
to find the correct lens. Nevertheless, our default parameter sits in a
good middle-ground. Both \TFTOpt{} and \NTFTOpt{} synthesized the correct lens
less often than \ZTOpt.  \TFTOpt{} demands too good a lens, so the search
procedure keeps spinning, while \NTFTOpt{} terminates early, not
always trying hard enough to find the correct lens.
% end evaluation
